Do return series have power-law tails with the same index?

### 2017-05-02 papers/FX/draft.tex
Completed the proof about monotonicity.

data from 1 January 2010 to 31 December 2014.

@@

\subsection{An utility based argument for equal tail indices}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$.
  Here $v$ can be thought of as the {\em certainty payoff} equivalent to the
 risky payoff $C$; $\delta$ tunes the {\em threshold of 
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappoinment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
\par
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:u}
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>-1\,. 
\eeam
%In practice it is more often to observe $\gamma > 1$ {\blue Let's
%consult Casper for references.} Therefore we adopt a more convenient
For the sake of argument, we assume that an investor initially has one
unit of wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
bond with interest rate $r>0$ and $\phi$ units in a risky asset with
return $X$ over one time unit, i.e.,
\beam\label{eq:xxie1}
  C = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
\eeam
Then we have 
\beao
\wt u(\phi,F_X) &=& \E [u(C)] + b\, \E \big[u(C)\1(C \le  \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
\eeao
where $F_X$ is the \df\ of $X$ with density $f_X$, and
\beao
  q = %q(r, \phi) &:=& \log \left( {
      %\delta v - (1 - \phi) e^r
      %\over
      %\phi
    % \right) 
\log\Big(
    \ex^r + {\delta \,v - \ex^r \over \phi}
  \Big)\,.
\eeao
Note that $C \le \delta v$ \fif\ $X \le  q$. 
\par
Naturally, if an agent invests in
a risky asset instead of a riskless bond, he expects to obtain a
higher (on average) return from the risky asset than he is guaranteed from the
riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
$q > r $. For given $b,\delta,v$, $\wt u$ is only a functional of $\phi$ and $F_X$,
$\wt u=\wt u (F_X, \phi)$.
We assume that
\beao
\wt u_{\rm max}=\wt u_{\rm max}(F_X) = \max_{0 \le  \phi \leq 1} \wt u (F_X,\phi)
\eeao
exists.

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
we assume the following toy model. 
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where {\red where needed?}
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(\phi,F_X)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x \le  q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
\par
We observe the following property which also applies to \eqref{eq:xxie1.0}. 
\ble\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\ele
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure~\ref{fig:sectors_parameters}.
\par
We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto}
for the power-utility \fct\ \eqref{eq:u}.
A particularly interesting special case is when the left and the right
tails have equal parameters, i.e. when $\alpha = \beta$ and $K =
K'$. In this case \eqref{eq:pareto} becomes
\begin{eqnarray*}
  \wt u(\phi, F_X) &=& \alpha p K^\alpha (1 + b)
  \int_{0}^\infty {
    u(C(x)) \left[
      1 - {b \over 1 + b}\1_{\{x > q\}}
      \right]
    + u(C(-x))
    \over
    (K + x)^{\alpha + 1}
  } dx
  - b u(\delta v) F_X(q)
\end{eqnarray*}
A detailed analysis of the montonicity of
\[
u(C(x)) \left[
  1 - {b \over 1 + b}\1_{\{x > q\}}
  \right]
+ u(C(-x))
\]
which comprises lemma \ref{lemma:III}, allows us to apply lemma
\ref{lemma:I} and conclude the following
\begin{itemize}
\item I am a good guy
\item So are you
\end{itemize}

For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
quantity is denoted by $I(\alpha, K)$.
\par
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
This is shown in 
Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
 \caption{The graphs show $\hat\phi$, the optimal equity allocation
    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
    the utility \fct\ $u$ from \eqref{eq:u}
    for $\xi = 1/2$ (left) and $\xi = 4$
    (right).
  }
  \label{fig:phi_hat_pareto}
\end{figure}

@@

\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto5e-1.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto4.pdf}
  \end{minipage}
  \caption{The quantities $I(\alpha, K)$ for the power-utility \fct\ with $\xi=1/2$ (left) and $\xi=4$ (right).
The graphs clearly show that  $I(\alpha, K)$ increases with $\alpha$ and decreases with $K$.
  }
  \label{fig:preference_pareto}
\end{figure}

@@

Let $u(\cdot)$  and $C(\cdot)$ be defined as in \eqref{eq:u} and \eqref{eq:xxie1} respectively. Define
U(x) =
  1 - {b \over 1 + b}\1_{\{x > q\}}

@@

x \geq 0

@@

\item  Hello

@@

  It is convenient to re-write $U(x)$ as

@@

  U(x) = 
  -{\hat\phi^{-\xi} \over \xi} \left\{

@@

    \left(1 - {b \over 1 + b}\1_{\{x > q\}}\right)
    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^x\right]^{-\xi}

@@

    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^{-x}\right]^{-\xi}
  }_{U'(x)}

@@

  Thus the monotonicity of $U(x)$ is the opposite of $U'(x)$. For
  convenience of writing, let $a = (1 - \hat\phi)e^r/\hat\phi$.
  First of all, we find the conditions on which $U'(x)$ is monotonic.

@@

    {\pd U'(x) \over \pd x}

@@

    x \neq q

@@

  ${\pd U'(x) \over \pd x} = 0$ is equivalent to

@@

    {a + y \over a y + 1} &=& 
    \left( \1_{\{y \leq e^q\}} + c \1_{\{y > e^q\}} \right)
    y^{1 -\xi \over 1+\xi} \\
    \underbrace{\log(a + y) - \log(a y + 1)}_{f(y)} &=& \underbrace{{1
        -\xi \over 1+\xi} \log(y)
      + \log(c)\1_{\{y > e^q\}} }_{g(y)} 

@@

  where we have defined $y = e^x$, $c = (1 + b)^{-1/(1 + \xi)}$ and
    f(y) &=& \log(a + y) - \log(a y + 1) \\
    g(y) &=& {1 -\xi \over 1+\xi} \log(y) + \log(c) \1_{\{y > e^q\}}
  Observe
    f(y) - g(y) &=& \log\left(
      {a + y \over a y + 1}
      y^{\xi - 1 \over \xi + 1}
    \right) - \log(c)\1_{\{y > e^q\}} \label{eq:f-g}
  and when $y \neq e^q$,
    {\pd f(y) \over \pd y} &=& {1 - a^2 \over (a + y) (a y + 1)}
    \label{eq:fd}\\
    {\pd g(y) \over \pd y} &=& {1 - \xi \over (1 + \xi) y}
    \label{eq:gd}
  Furthermore

@@

    y \neq e^q

@@

    real, ${\pd (f -g)(y) \over \pd y} = 0$ has no solution on $(1,
    \infty)\setminus \{e^q\}$; $(f -g)(y)$ is monotonic on both
    $(1, e^q)$ and $(e^q, \infty)$.

@@

      is monotonically increasing on both
      $(1, e^q)$ and $(e^q, \infty)$, because the coefficient of the
      $y^2$ term of the numerator of \eqref{eq:f-g-der}, i.e. $a (\xi
      - 1)$ is positive.
      \item If $f(e^q) > g(e^q)$, then clearly $f(y) = g(y)$ has no
        solution on $(e^q, \infty)$; we also note $f(1) = g(1) =
        0$. Thus $f(y) = g(y)$ has no solution on $(1, e^q)$ either.
        Now it can be concluded that $U'(\cdot)$ is monotonic on
        $(1, q)$ and $(q, \infty)$. To determine the monotonicity of
        $U'(\cdot)$, it remains to check the signs of
        ${\pd U'(x) \over \pd x}$  on  $(0, q)$ and on $(q, \infty)$
        as well as the values of $\lim_{x \uparrow q} U'(x)$ and
        $\lim_{x \downarrow q} U'(x)$. Obviously

@@

          \left.{\pd U'(x) \over \pd x}\right|_{x=0} &=& 0 \\

@@

        For a small $\epsilon > 0$, the sign of ${\pd U'(x) \over \pd

@@

        $\left.{\pd^2 U'(x) \over \pd x^2}\right|_{x=0}$:

@@

        \left.{\pd^2 U'(x) \over \pd x^2}\right|_{x=0} =

@@

      % conversely, $f(e^q) < g(e^q)$ implies $f(y) = g(y)$ must have a
      % solution on $(e^q, \infty)$, since \eqref{eq:f-g} tells us
      % $(f-g)(y) \to \infty$ as $y \to \infty$.

    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, $(f - g)(y)$
      is monotonically decreasing.
    \end{enumerate}

  % \item If $a < \min\{1, \xi\}$ or $a > \max\{1, \xi\})$, ${\pd (f
  %     -g)(y) \over \pd y} = 0$ has 2 real solutions.
  % \end{enumerate}

  % \begin{enumerate}
  % \item If $a < 1$ and $\xi > 1$
  %   \begin{enumerate}
  %   \item when $x > q$, $U'(x)$ being monotonic is equivalent to ${\pd U'(x)
  %       \over \pd x}$ having no zero on $(q, \infty)$.

  %     \begin{enumerate}
  %     \item If $f (e^q) > g(e^q)$ it is clear $f(y) = g(y)$
  %       has no solution on $(e^q, \infty)$.
  %     \item If $f (e^q) <  g(e^q)$ it is also clear $f(y) = g(y)$ must have
  %       one solution on $(e^q, \infty)$.
  %     \end{enumerate}
  %   \end{enumerate}

  %   \item If $a > 1$ and $\xi < 1$
  %     \begin{enumerate}
  %     \item when $x > q$, the same calculation that leads to
  %       \eqref{eq:fd} and \eqref{eq:gd} gives ${\pd f(y) \over \pd y}
  %       < 0$ and ${\pd g(y) \over \pd y} > 0$. So
  %     \begin{enumerate}
  %     \item If $f (e^q) >  g(e^q)$, $f(y) = g(y)$ must have
  %       one solution on $(e^q, \infty)$.
  %     \item If $f (e^q) < g(e^q)$, $f(y) = g(y)$ has no solution on $(e^q, \infty)$.
  %     \end{enumerate}
  %     \end{enumerate}

  %     \item If $a < 1$ and $\xi < 1$
  %       \begin{enumerate}
  %       \item when $x > q$
  %         \begin{enumerate}
  %         \item If $f (e^q) -  g(e^q) > 0$, we observe $f(y) - g(y)
  %           \to -\infty$ as $y \to \infty$. So by continuity $f(y) -
  %           g(y) = 0$ must have at least one solution on $(e^q, \infty)$.
  %         \end{enumerate}
  %       \end{enumerate}
  %     \item If $a > 1$ and $\xi > 1$
  %       \begin{enumerate}
  %       \item when $x > q$, both $f(y)$ and $g(y)$ decreases. Observe
  %         In this case, if $f(e^q) > g(e^q)$, there is no solution 
  %       \end{enumerate}
  \end{enumerate}
\end{proof}

>>>>

data from 4th January 2010 to 31 December 2014.

@@

\subsection{An utility based argument for equal tail indices}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$.
  Here $v$ can be thought of as the {\em certainty payoff} equivalent to the
 risky payoff $C$; $\delta$ tunes the {\em threshold of 
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappoinment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
\par
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:u}
u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
\eeam
In practice it is more often to observe $\gamma > 1$ {\blue Let's
consult Casper for references.} Therefore we adopt a more convenient
notation in this case
\beao
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
\eeao

%In practice it is more often to observe $\gamma > 1$ {\blue Let's
%consult Casper for references.} Therefore we adopt a more convenient
For the sake of argument, we assume that an investor initially has one
unit of wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
bond with interest rate $r>0$ and $\phi$ units in a risky asset with
return $X$ over one time unit, i.e.,
\beam\label{eq:xxie1}
  C = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
\eeam
Then we have 
\beao
\wt u(\phi,F_X) &=& \E [u(C)] + b\, \E \big[u(C)\1(C \le  \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
\eeao
where $F_X$ is the \df\ of $X$ with density $f_X$, and
\beao
  q = %q(r, \phi) &:=& \log \left( {
      %\delta v - (1 - \phi) e^r
      %\over
      %\phi
    % \right) 
\log\Big(
    \ex^r + {\delta \,v - \ex^r \over \phi}
  \Big)\,.
\eeao
Note that $C \le \delta v$ \fif\ $X \le  q$. 
\par
Naturally, if an agent invests in
a risky asset instead of a riskless bond, he expects to obtain a
higher (on average) return from the risky asset than he is guaranteed from the
riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
$q > r $. For given $b,\delta,v$, $\wt u$ is only a functional of $\phi$ and $F_X$,
$\wt u=\wt u (F_X, \phi)$.
We assume that
\beao
\wt u_{\rm max}=\wt u_{\rm max}(F_X) = \max_{0 \le  \phi \leq 1} \wt u (F_X,\phi)
\eeao
exists.

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
we assume the following toy model. 
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
where
$0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(\phi,F_X)\nonumber
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x <  q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
We observe the following property which also applies to \eqref{eq:xxie1.0}. 
\ble\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\ele
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto}
for the power-utility \fct\ \eqref{eq:u}.
A particularly interesting special case is when the left and the right
tails have equal parameters, i.e. when $\alpha = \beta$ and $K =
K'$. In this case \eqref{eq:pareto} becomes
  \wt u(\phi, F_X) &=& \alpha p K^\alpha (1 + b)
  \int_{0}^\infty {
      1 - {b \over 1 + b}\1_{\{x \ge q\}}
    (K + x)^{\alpha + 1}
  } dx
  - b u(\delta v) F_X(q)
  \label{eq:ipf}
First of all we note that the function
U_{\text{all}} =
  1 - {b \over 1 + b}\1_{\{x \ge q\}}
is not monotonic. To see this we may re-write the above as
U_{\text{all}} =
  -{\phi^{-\xi} \over \xi} \left\{
    \left(1 - {b \over 1 + b}\1_{\{x \ge q\}}\right)
    \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}
    \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
  }_{U(x)} \right\}
  \label{eq:frg}
Direct computation gives
  {\pd U(x) \over \pd x}
  \quad x \neq q
where $a = (1 - \phi) e^r/\phi$. For $x \ge q$, we have
  {1 \over \xi}
  {\pd U(x) \over \pd x} &=& {x' \over (a + x')^{\xi + 1}}
  - {x'^\xi \over (a x' + 1)^{\xi + 1} (1 + b)}
where we have substituted $x'$ for $e^{-x}$. When $x$ is very large
i.e. when $x'$ is very close to 0, it is useful to expand the last
expression at $x' = 0$. This yields
  {1 \over \xi}
  {\pd U(x) \over \pd x}
  = {x' \over a^{\xi + 1}} + o(x')
So we see that when $x$ is sufficiently large
${\pd U(x) \over \pd x} > 0$. On the other hand, at the discontinuity
point $x = q$, we have
\lim_{x \downarrow 0} \left(
  U(q + x) - U(q - x)
= - {
  (a + e^q)^{-\xi} b
  1+ b
} < 0
So $U(x)$ and hence \eqref{eq:frg} is not monotonic. we resort to
numerical methods to gain some understanding of how $\tilde u(F_X,
\phi)$ changes with $\alpha$.
For this reason, we replace $\phi$ in the first integral in
\eqref{eq:pp} by the optimal $\hat\phi$. The resulting quantity is
denoted by $I(\alpha, K)$.
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for
given values of $K, K', \alpha, \beta$.  This is shown in 
Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
 \caption{The graphs show $\hat\phi$, the optimal equity allocation
    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
    the utility \fct\ $u$ from \eqref{eq:u}
    for $\xi = 1/2$ (left) and $\xi = 4$
    (right).
  \label{fig:phi_hat_pareto}

@@

The following graph (figure \ref{fig:preference_pareto}) shows how
$\tilde u(\alpha, K, \wt\phi)$ changes with $\alpha$ and $K$.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto5e-1.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{preference_pareto4.pdf}
  \end{minipage}
  \caption{The quantities $I(\alpha, K)$ for the power-utility \fct\
    with $\xi=1/2$ (left) and $\xi=4$ (right). The graphs clearly show
    that  $I(\alpha, K)$ increases with $\alpha$ and decreases with
    $K$.
  }
  \label{fig:preference_pareto}
\end{figure}

When the parameter $b$ is very small, a GDA preference is closely
approximated by a mean-utility preference, which GDA reduces to when
$b = 0$. As shown in the proof of lemma \ref{lemma:III}, the function
$U_{\text{all}}$ is monotonically increasing or decreasing when $\xi$
and $(1 - \phi) e^r / \phi$ satisfy certain conditions. Moreover, it
is also easily checked that, when $x \in (0, K(e^{1/\alpha} - 1)]$,
the density function in the integral of \eqref{eq:ipf} is increasing
with $\alpha$; when $x \in (K(e^{1/\alpha} - 1), \infty)$, it is
decreasing with $\alpha$. So, following the same line of arguments
using lemma \ref{lemma:I} as in the proof of lemma \ref{thrm:I},
it can be seen $\tilde u_{\text{max}}(\alpha, K)$ is
increasing/decreasing with $\alpha$ when $U_{\text{all}}$ is
decreasing/increasing.

@@

Let $u(\cdot)$  and $C(\cdot)$ be defined as in \eqref{eq:u} and
\eqref{eq:xxie1} respectively. Define
U_{\text{all}} &=& u(C(x))

@@

x \geq 0 \\
a &=& {
  (1 - \phi) e^r
  \phi
} \\
The following holds true:

@@

\item If $\max\{a, 1\} < \xi$, $U_{\text{all}}(\cdot)$ is
  monotonically decreasing.
\item If $a < \xi < 1$ and $(a + y_-)/(a y_- + 1) <
  y_-^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotonically
  decreasing.
\item If $\xi < a < 1$, $U_{\text{all}}(\cdot)$ is monotonically
  increasing.
\item If $1 < \xi < a$ and $(a + y_+)/(a y_+ + 1) >
  y_+^{(1-\xi)/(1+\xi)}$, $U_{\text{all}}(\cdot)$ is monotonically
  increasing.
\item In other case, $U_{\text{all}}(\cdot)$ is not monotonic.

@@

  It is convenient to re-write $U_{\text{all}}$ as

@@

  U_{\text{all}} =
  -{\phi^{-\xi} \over \xi} \left\{

@@

      \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}

@@

      \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
    }_{U(x)}

@@

  Thus the monotonicity of $U_{\text{all}}(x)$ is the opposite of $U(x)$. For
  convenience of writing, let $a = (1 - \phi)e^r/\phi$.
  First of all, we find the conditions on which $U(x)$ is monotonic.

@@

    {\pd U(x) \over \pd x}

@@

@@

  ${\pd U(x) \over \pd x} = 0$ is equivalent to

@@

    {a + y \over a y + 1} &=& y^{1 -\xi \over 1+\xi} \\
    \underbrace{\log(a + y) - \log(a y + 1)}_{f(y)} &=&
    {1 -\xi \over 1+\xi} \log(y)

@@

  where we have defined $y = e^x$, and $f(y)$, $g(y)$ as above. Observe

@@

@@

    real, ${\pd (f -g)(y) \over \pd y} = 0$ has no solution on
    $(1, \infty)$; $(f -g)(y)$ is monotonic on $(1, \infty)$.

@@

      is monotonically increasing on $(1, \infty)$, because the
      coefficient of the $y^2$ term of the numerator of
      \eqref{eq:f-g-der}, i.e. $a (\xi - 1)$ is positive. We note
      $f(1) = g(1) = 0$; thus on $(1, \infty)$, there is no solution
      to $f(y) = g(y)$. It can be concluded that $U(\cdot)$ is
      monotonic on $(0, \infty)$. Furthermore

@@

          \left.{\pd U(x) \over \pd x}\right|_{x=0} &=& 0 \\

@@

        For a small $\epsilon > 0$, the sign of ${\pd U(x) \over \pd

@@

        $\left.{\pd^2 U(x) \over \pd x^2}\right|_{x=0}$:

@@

        \left.{\pd^2 U(x) \over \pd x^2}\right|_{x=0} =

@@

        Thus ${\pd U(x) \over \pd x} > 0$ for $x \in (0, \infty)$;
        $U_{\text{all}}$ is monotonically decreasing.
    \item If instead $\xi < 1$, i.e. $\xi < a < 1$, by a similar
      argument as in the previous case, $U_{\text{all}}$ is monotonically
      increasing on $(0, \infty)$.
  \item If $a < \min\{1, \xi\}$ and $\xi > 1$, i.e. $a < 1 < \xi$, it
    is clear
    y_- = {
      a^2 - \xi - \sqrt{(a^2 - 1) (a^2 - \xi^2)}
    } < 0
    It remains to compare $y_+$ with $1$ to determine whether
    ${\pd (f -g)(y) \over \pd y} = 0$ has a solution on $(1, \infty)$.
    Assume $y_+ > 1$. Then
      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1)
      \nonumber \\
      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0 \label{eq:mi6}
    This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotonically increasing
    on $(1, \infty)$. Following the same analysis as in the case
    (1.a), one can see $U_{\text{all}}$ is monotonically decreasing.

  \item If $a < \min\{1, \xi\}$ and $\xi < 1$, i.e. $a < \xi < 1$, it
    is clear $y_- > 0$ and $y_+ < y_-$. Moreover, $y_- > 1$ is equivalent
    to
    \[
    \sqrt{(a^2 - 1) (a^2 - \xi^2)} > a^2 - \xi a + a - \xi
    = (a + 1) (a - \xi)
    \]
    The last inequality is obviously true in this case. So $y_- >
    1$. ${\pd (f-g)(y) \over \pd y}$ has a maximum at
    $y_- \in (1, \infty)$. We know $(f - g)(y_-)$ is a maximum because
    \eqref{eq:f-g-der} shows that, for $y > y_-$,
    ${\pd (f-g)(y) \over \pd y} < 0$.

    If $(f-g)(y_-) < 0$, $(f-g)(y) = 0$ has no solution on $(1,
    \infty)$. Hence $U(\cdot)$ is monotonic on both $(0, \infty)$.
    The same analysis as in the case (1.a) shows $U_{\text{all}}$
    is monotonically decreasing on $(0, \infty)$.

    If $(f-g)(y_-) > 0$, $(f-g)(y) = 0$ must have a solution on $(y_-,
    \infty)$. So $U(\cdot)$ is not monotonic.

    \item $a > \max\{1, \xi\}$ and $\xi > 1$, i.e. $1 < \xi < a$. By
      the same argument that leads to \eqref{eq:mi6}, we see
      $y_+ > 1$. From \eqref{eq:f-g-der} it is clear $(f-g)(y_+)$ is a
      minimum. If $(f-g)(y_+) > 0$, there is no solution to
      $(f-g)(y) = 0$ on $(1, \infty)$; $U(x)$ is monotonic. By the
      same analysis as in the case (1.a), we know $U(x)$ is
      monotonically decreasing and $U_{\text{all}}$ is monotonically increasing.

      if $(f-g)(y_+) < 0$, there must be a solution to $(f-g)(y) = 0$
      on $(y_+, \infty)$. $U_{\text{all}}$ is not monotonic.
  \end{enumerate}
\end{proof}
